---
title: "Assignment_5"
author: "Bhupesh"
date: "17 October 2016"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

################################################################################
# GARCH/ ARCH model
################################################################################

```{r, warning=FALSE}

# load the required packages

library(xts)
library(PerformanceAnalytics)
library(quantmod)
library(rugarch)
library(car)
library(FinTS)
library(forecast)

Data<-read.csv("nyse.csv",sep=",",header=TRUE)
temp<-as.Date(Data$ENTRY,format="%Y:%m:%d")
head(temp)
tail(temp)

```

# Extract adjusted closing prices

```{r, warning=FALSE}
nys <- Data[, "RATE", drop=F]
nys<-xts(nys,temp) # define nys as xts object



# Remove first NA observation
nys.r <- nys[-1,]
colnames(nys.r) <- "nyse returns"

```

# plot returns

```{r}
plot(nys.r)

plot(density(nys.r), main="Actual Returns")

```

# plot returns with squared and absolute returns

```{r, warning=FALSE}
dataToPlot <- cbind(nys.r, nys.r^2)
colnames(dataToPlot) <- c("Returns", "Returns^2")
plot.zoo(dataToPlot, main="nyse Daily Returns", col="blue")


par(mfrow=c(2,1))
a1<-Acf(nys.r, main="nyse Returns")
a2<-Acf(nys.r^2, main="nyse Returns^2")
par(mfrow=c(1,1))

par(mfrow=c(2,1))
pacf(nys.r, main="nyse Returns")
pacf(nys.r^2, main="nyse Returns^2")
par(mfrow=c(1,1))

table.Stats(nys.r)

arima(nys.r, order=c(2,0,0)) 
arima(nys.r, order=c(2,0,1))   
arima(nys.r, order=c(1,0,0))   
arima(nys.r, order=c(1,0,1))
arima(nys.r, order=c(2,0,2))
arima(nys.r, order=c(0,0,1))
arima(nys.r, order=c(0,0,0))
arima(nys.r, order=c(3,0,1))
arima(nys.r, order=c(3,0,0))
arima(nys.r, order=c(2,0,2))
arima(nys.r, order=c(1,0,2))


fit<-arima(nys.r, order=c(2,0,0)) 
fit

Ajb <- matrix(nrow=30, ncol=2)
for(i in 1:30){
  jb <- Box.test(fit$residuals^2, lag=i, type="Ljung-Box")
  round(Ajb[i,1] <- jb$statistic,2)
  round(Ajb[i,2] <- jb$p.value,2)
}
Ajb <- round(data.frame(Ajb),2)
dimnames(Ajb) <- list(paste("Lag", 1:30, sep="."), c("stat.",
                                                     "pval"))
Ajb

```

# use ArchTest() function from FinTS package for ARCH LM test

```{r, warning=FALSE}
ArchTest(fit$residuals^2,lags=1)
ArchTest(fit$residuals^2,lags=4)

plot(fit$residuals^2)

par(mfrow=c(2,1))
acf(fit$residuals^2,lag.max = 100)
pacf(fit$residuals^2,lag.max = 100)
par(mfrow=c(1,1))

```

#GARCH(1,0) MODEL

```{r, warning=FALSE}
g10.s <- ugarchspec(variance.model = list(garchOrder=c(1,0)), 
                    mean.model = list(armaOrder=c(2,0)))
# Once specifined you may use ugarchfit() to estimate
nyse.g10.fit1 <- ugarchfit(spec=g10.s, data=nys.r,
                           solver.control=list(trace = 1))                          
nyse.g10.fit1

```

#GARCH(0,1) MODEL

```{r, warning=FALSE}
g01.s <- ugarchspec(variance.model = list(garchOrder=c(0,1)), 
                    mean.model = list(armaOrder=c(2,0)))
# Once specifined you may use ugarchfit() to estimate
nyse.g01.fit1 <- ugarchfit(spec=g01.s, data=nys.r,
                           solver.control=list(trace = 1))                          
nyse.g01.fit1

```

#GARCH(1,1) MODEL

```{r, warning=FALSE}
g11.s <- ugarchspec(variance.model = list(garchOrder=c(1,1)), 
                    mean.model = list(armaOrder=c(2,0)))
nyse.g11.fit1 <- ugarchfit(spec=g11.s, data=nys.r,
                         solver.control=list(trace = 1))                          
nyse.g11.fit1


```

#GARCH(1,1) Model with t distribution

```{r, warning=FALSE}
g11.t <- ugarchspec(variance.model = list(garchOrder=c(1,1)), 
                    mean.model = list(armaOrder=c(2,0)),distribution.model = "std")
nyse.g11.fit2 <- ugarchfit(spec=g11.t, data=nys.r,
                          solver.control=list(trace = 1))                          
nyse.g11.fit2

```

#IGARCH(1,1) Model

```{r}
i11.s <- ugarchspec(variance.model = list(model="iGARCH",garchOrder=c(1,1)), 
                    mean.model = list(armaOrder=c(2,0)))
nyse.i11.fit <- ugarchfit(spec=i11.s, data=nys.r,
                          solver.control=list(trace = 1))  
nyse.i11.fit

```

# Residuals

```{r, warning=FALSE}
#residuals:ge(t)
res0 <- residuals(nyse.g11.fit1)
head(res0)
# standard residuals: gv(t)
sres0 <- residuals(nyse.g11.fit1, standardize =TRUE)
head(sres0)

#residuals: ie(t)
res1 <- residuals(nyse.i11.fit)
head(res1)
# standard residuals: iv(t)
sres1 <- residuals(nyse.i11.fit, standardize =TRUE)

head(sres1)
```

# Residual plots

```{r, warning=FALSE}
resToPlot <-cbind(res1,sres1,res0,sres0)
colnames(resToPlot) <- c("ie(t)", "iv(t)","ge(t)","gv(t)")
plot.zoo(resToPlot,col="blue")

Box.test(sres1,lag=5, type=c("Ljung-Box"),fitdf=0)
Box.test(sres1,lag=10, type=c("Ljung-Box"),fitdf=0)
Box.test(sres1,lag=15, type=c("Ljung-Box"),fitdf=0)

Box.test(sres0,lag=5, type=c("Ljung-Box"),fitdf=0)
Box.test(sres0,lag=10, type=c("Ljung-Box"),fitdf=0)
Box.test(sres0,lag=15, type=c("Ljung-Box"),fitdf=0)

# plot autocorrelations of standardised residuals for GARCH(1,1)
par(mfrow=c(2,1))
Acf(sres0,lag.max =10, main="std. residuals") 
Acf(sres0^2,lag.max =10, main="std. residuals^2")
par(mfrow=c(1,1)) 

par(mfrow=c(2,1))
Pacf(sres0,lag.max =10, main="std. residuals")
Pacf(sres0^2,lag.max =10, main="std. residuals^2")
par(mfrow=c(1,1))

# plot autocorrelations of standardised residuals for IGARCH(1,1)
par(mfrow=c(2,1))
Acf(sres1,lag.max =10, main="std. residuals") 
Acf(sres1^2,lag.max =10, main="std. residuals^2")
par(mfrow=c(1,1)) 

par(mfrow=c(2,1))
Pacf(sres1,lag.max =10, main="std. residuals")
Pacf(sres1^2,lag.max =10, main="std. residuals^2")
par(mfrow=c(1,1))

```
 
 In addition the sign bias test show evidence for leverage effect. hence an EGARCH 
 or TARCH may give better results. 

```{r, warning=FALSE}
tgarch.spec = ugarchspec(variance.model=list(model="fGARCH",garchOrder=c(1,1),submodel="TGARCH"),
                          mean.model=list(armaOrder=c(2,0)))
nyse.tgarch.fit = ugarchfit(tgarch.spec, nys.r)
nyse.tgarch.fit
```

# Nelson's egarch model

```{r, warning=FALSE}
egarch1.spec = ugarchspec(variance.model=list(model="eGARCH",
                                               garchOrder=c(1,1)),
                           mean.model=list(armaOrder=c(2,0)))
nyse.egarch1.fit = ugarchfit(egarch1.spec, nys.r)
nyse.egarch1.fit
```

# GJR garch model

```{r, warning=FALSE}
gjrgarch.spec = ugarchspec(variance.model=list(model="gjrGARCH",
                                                 garchOrder=c(1,1)),
                             mean.model=list(armaOrder=c(2,0)))
nyse.gjrgarch.fit = ugarchfit(gjrgarch.spec, nys.r)
nyse.gjrgarch.fit
```

# Compare information criteria and choose the best among all

```{r, warning=FALSE}
model.list = list(garch = nyse.g11.fit1,
                  egarch = nyse.egarch1.fit,
                  igarch=nyse.i11.fit,
                  gjrgarch = nyse.gjrgarch.fit)
info.mat = sapply(model.list, infocriteria)
rownames(info.mat) = rownames(infocriteria(nyse.g11.fit1))
info.mat


```

