---
title: "Assignment_2_ARMA"
author: "Bhupesh"
date: "19 August 2016"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Fitting the ARMA model to the given data set

#Step 1: Import the data set

```{r}
inflation1 <- read.csv("~/Applied_Trix/inflation1.csv")
head(inflation1)
```

#Step 2: Convert the data set to a time series

```{r}
ts_inflation_f1<-ts(inflation1$Inf.)

```

#Step 3: Import the file to provided to draw graphs and load the required library

```{r}
options(digits=4, width=70)
source("E:/Study/Semester 3/macro & fin ecotrix/graphfun.R")
library(ggplot2)
library(grid)
library(gridExtra)
```

#Step 4: Plot the time series

```{r}

ts.graph1(ts_inflation_f1)

```

#Step 5: Plot sample ACF and sample PACF using the function

```{r}

ts.graph4(ts_inflation_f1,30)

```

It is clear from smaple ACF that there is a significiant lag of 2 periods so this can either be a AR(2) process or an ARMA(2,1) process. Let us estimate the ACI for these process

# Step 6: Estimate the cofficient

```{r}
#for AR(2) process
arima(ts_inflation_f1, order = c(2,0,0))

#for ARMA(2,1)
arima(ts_inflation_f1, order = c(2,0,1))


```

# Step 7: Diagnostic

Since ACI for ARMA(2,1) is less than AR(2) process therefore ARMA(2,1) is a better fit model for the given data.

The cofficient are for y(t-1)=1.684, for y(t-2)=-0.76, for E(t-1)=-0.844 and constant = 5.337
